\Huge \bf Implementation of a quadrotor UAV

### 2012-07-14 latex/paper/paper.tex
+ detalles

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

@@

A quadcopter platform can be imagined performing countless tasks. Many applications are being developed based on such platforms, but how do they fly? A aerial vehicle is inherently unstable, staying still is not a simple task. This paper focuses on the stabilization of a quadcopter, explaining how to work with an IMU, the development of the mathematical model that represents the system, the filtering techniques applied to sensor data, and the control system that will allows to quadcopter to fly.

The platform is based on a commercial quadcopter, the frame and the motor control system were preserved, but the IMU and intelligence were replaced by the flight controller that was developed. An ARM-Cortex-A8 processor running Linux performs the computations requiered to convert raw IMU data and combine it using an EKF, overcoming the noise generated by the vibrations of the motors, and providing a reliable estimation of the state variables. Once the current state is known, the LQR algorithm is used to derive control actions required to bring the system to the desired setpoint.

@@

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{./pics_paper/quad_coord.jpg}
	\caption{Model of a quadcopter}
	\label{fig:quad}
\end{figure}

@@

The equations that govern the dynamics of a quadcopter are bla.

@@

In order to perform adequate control actions, a reliable estimation of the state variables must be available in real time. A Kalman Filter uses the mathematical model for the system and the current estimation of the state vector to predict what should happen next, and corrects the prediction with the information read from the sensors, taking into consideration high much confidence is placed on the prediction and how much on the measurements. This prediction-correction technique allows to have a smooth state estimation, without the typical delay introduced by filtering. Even small delays can affect the performance of the system.

@@

The theory behind a standard Kalman Filter does not hold for a system that is not linear. The model for the quadcopter is highly non-linear, so an EKF is implemented based on \cite{bib:kalman} and \cite{bib:kalman2}. An EKF is not optimal, and the performance is highly dependant on the linearizacion \cite{bib:kay}. The performance achieved is considered acceptable.

@@

@@

The software is spread across several boards: An ARM-Cortex-A8 on the flight controller, an ATMega328p on the IMU, and four microprocessors\footnote{C8051F330/1 - Silicon Labs.} on the ESCs, one for each motor. The code running on the ESCs is not available, but the communication protocol was decoded using a logic analyzer. The firmware on the IMU takes care of reading from all the sensors (using $I^2C$) and sending a new frame of data every 10ms over a UART. The code running on the flight controller processes the data from the IMU, calculates the necessary control actions, and sets the desired speed for each motor by sending the appropriate commands, via $I^2C$, to the ESCs.

>>>>

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

@@

A quadcopter platform can be imagined performing countless tasks. Many applications are being developed based on such platforms, but how do they fly? A aerial vehicle is inherently unstable, staying still is not a simple task. This paper focuses on the stabilization of a quadcopter, explaining how to work with an IMU\cite{bib:calib_imu}\cite{bib:bola}\cite{bib:alain}, the development of the mathematical model \cite{bib:quadrotor-bible}\cite{bib:curso_quad} that represents the system, the filtering techniques applied to sensor data\cite{bib:kalman}\cite{bib:kalman2}, and the control system\cite{bib:quadrotor-bible}\cite{bib:hakas} that will allows to quadcopter to fly.

The platform is based on a commercial quadcopter, the frame and the motor control system were preserved, but the IMU and intelligence were replaced by the flight controller that was developed. An ARM-Cortex-A8 processor running Linux performs the computations requiered to convert raw IMU data and combine it using an EKF, overcoming the noise generated by the vibrations of the motors, and providing a reliable estimation of the state variables. Once the current state is known, the LQR algorithm\cite{bib:lqrnotes}\cite{bib:lqr-discreto} is used to derive control actions required to bring the system to the desired setpoint.

@@

\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{./pics_paper/quad_coord.jpg}
	\caption{Model of a quadcopter - Intertial reference (blue), Non-Inertial reference (red)}
	\label{fig:quad}
\end{figure}

@@

The equations that govern the dynamics of a quadcopter are presented in equation \ref{eq:modelo}. The reference frames used are shown in figure \ref{}.

@@

In order to perform adequate control actions, a reliable estimation of the state variables must be performed in real time. A Kalman Filter achieves this using the mathematical model for the system to predict what should happen given the current state, and corrects the prediction with the information read from the sensors, taking into consideration how much confidence is placed on the prediction and how much on the measurements. This weighted prediction-correction technique allows a smooth state estimation, without the typical delay introduced by filtering. Even small delays can affect the performance of the system.

@@

%TODO pasar figura a ingl\'e
\begin{figure}
	\centering
	\includegraphics[width=1\columnwidth]{../pics_kalman/diagrama_kalman_2.png}
	\caption{}
	\label{fig:diagrama_kalman}
\end{figure}

The theory behind a standard Kalman Filter does not hold for a system that is not linear. The model for the quadcopter is highly non-linear, so an EKF is implemented based on \cite{bib:kalman} and \cite{bib:kalman2}. An EKF is not optimal, and the performance is highly dependant on the linearizacion \cite{bib:kay}.

@@

The EKF proved to play a critical part. For example, the data provided by the accelerometer is unusable without adequate filtering, and a simple low pass filter (LPF) was implemented to reduce noise. The 60ms delay introduced by the LPF severly deteriorated the performance of the system. When the EKF was assigned the task of reducing noise, the perfomance was significatively improved.

@@

The software runs on several independant boards: An ARM-Cortex-A8 on the flight controller, an ATMega328p on the IMU, and four microprocessors\footnote{C8051F330/1 - Silicon Labs.} on the ESCs, one for each motor. The code running on the ESCs is not available, but the communication protocol was decoded using a logic analyzer. The firmware on the IMU takes care of reading from all the sensors (using $I^2C$) and sending a new frame of data every 10ms over a UART. The code running on the flight controller processes the data from the IMU, calculates the necessary control actions, and sets the desired speed for each motor by sending the appropriate commands, via $I^2C$, to the ESCs.
